Jury members' comments on the private PhD defense of \linebreak Gustavo Quintana-Carapia

### 2020-08-21 internal defense feedback/Answers-jury-members-report BK.tex
Second draft before public defense

	{\bfseries In Section 4.1.4, a discussion that compares IV methods with EIV methods has been added.}

@@

	{\bfseries The measurementy noise variance is prior information and in Chapter 5 we describe two ways of estimating $\sigma_{\bm{\epsilon}}$ In summary, the measurement noise can be extracted fom the sensor ressponse once the transient effects are smaller than the noise floor, or from the sensor response power spectrum. The online estimation of the measurement noise, simultaneously with the online step input estimation is a topic for future research.   }

>>>>

	{\bfseries In Section 4.1.4, a discussion that compares IV methods with EIV methods has beeen added.}

@@

	{\bfseries The measurementy noise variance is prior information and in Chapter 5 we describe two ways of estimating $\sigma_{\bm{\epsilon}}$ In summary, the measurement noise can be extracted fom the sensor ressponse once the transient effects are smaller than the noise floor, or from the sensor response power spectrum. The online estimation of the measurement noise, simultaneously with the online step input estimation is a topic for future reasearch.   }

### 2020-08-21 internal defense feedback/Answers-jury-members-report.tex
Second draft before public defense

    \renewcommand{\thefigure}{6.8}
    \begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterRampInput/fig/Fig_7.pdf} 
    \caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor's initial conditions are estimated with the ML method. After three iterations, \color{blue} the average of the relative error is smaller than 5 \% for each and everyone of the estimates. The estimate $\widehat{x}_{\mathrm{ini,2}}$ has the larger relative error near to 1 \%. \color{black} }
    \end{figure}

@@

	\color{blue}
	The normality assumption is necessary to provide more prior information into the method formulation.
    The second order Taylor series expansion (4.4) is developed considering the perturbation noise that gets into the regression matrix from the sensor transient response.
    Assuming only that the perturbation noise is distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$ is not enough.
    If we additionally assume normality for the perturbation noise, then we benefit from the knowledge of the third moment equal to zero due to symmetry, and the fourth moment equal to three times the squared variance, which can be disregarded since $3 \sigma_{\bm{\epsilon}}^4 \ll \sigma_{\bm{\epsilon}}^2$\color{black}.

@@

\begin{itemize}
	\item Two practical questions. On page 35-49 you consider a weight of 138.x grams. You sample relatively slowly (4 kHz) and after 500 samples we can consider that we have correct estimation. If I have a container that fills up and weight is constantly changing. Can we tackle this?
	\item  Consider an optical fiber link with a detector capable of 100 million measurements per second. Can I speed up the link by using your method? You have to do calculations: how high can you go in speed before running into troubles with DSP?
\end{itemize}

@@

    The errors-in-variables problem formulation (3.18) of the data-driven step input estimation method can be converted into an instrumental-variables problem.
    A guide to do this conversion is found in \cite{Soderstrom18} and in \cite{Pintelon12Book}. 
    The instrumental variable $\widetilde{\mathbf{Z}}$ has to be selected so that the normal equations (3.19) are expresed instead as
    \begin{equation} \tag{4.22} \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{y}} . \label{eqn:neq_siv} \end{equation}
    The selection of the instrumental variable has to be done in such a way that is correlated the regression matrix $\widetilde{\mathbf{K}}$, and uncorrelated with the perturbations $\bm{\epsilon}$.
    Assuming that the block-Hankel matrix $\widetilde{\mathbf{K}}$ is given by
    \begin{equation} \tag{4.23} \widetilde{\mathbf{K}} = \begin{bmatrix} G & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ G & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(n+1) & \Delta \widetilde{y}(n+2) & \cdots & \Delta \widetilde{y}(2n) \end{bmatrix} , \label{eqn:matrixK_r} \end{equation}
    then, the instrumental variable $\widetilde{\mathbf{Z}}$ should not contain the elements that construct $\widetilde{\mathbf{K}}$, i.e., the elements of the vector $\left( \widetilde{y}(1), \ldots, \widetilde{y}(2n+1) \right)^\top$.
    A possibility to generalize this idea is expressing the instrumental variable as
    \begin{equation} \tag{4.24} \widetilde{\mathbf{Z}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-4n) & \Delta \widetilde{y}(t-4n+1) & \cdots & \Delta \widetilde{y}(t-3n-1) \\ G & \Delta \widetilde{y}(t-4n+1) & \Delta \widetilde{y}(t-4n+2) & \cdots & \Delta \widetilde{y}(t-3n) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-3n) & \Delta \widetilde{y}(t-3n+1) & \cdots & \Delta \widetilde{y}(t-2n-1) \end{bmatrix} , \label{eqn:matrixZ_t} \end{equation}
    for a regression matrix
    \begin{equation} \tag{4.25} \widetilde{\mathbf{K}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-2n+1) & \Delta \widetilde{y}(t-2n+2) & \cdots & \Delta \widetilde{y}(t-n) \\ G & \Delta \widetilde{y}(t-2n+2) & \Delta \widetilde{y}(t-2n+3) & \cdots & \Delta \widetilde{y}(t-n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-n+1) & \Delta \widetilde{y}(t-n+2) & \cdots & \Delta \widetilde{y}(t) \end{bmatrix} , \label{eqn:matrixK_t} \end{equation}
    for $t = 4n+1, \ldots, N$.

@@

    \subsubsection*{4.1.4.2 Recursive generalized instrumental variables method}
    If we define
    \begin{equation} \tag{4.34} \begin{aligned} \mathbf{h} \left( \bm{\rho}, \bm{\theta} \right)  &=  \mathbf{r}_{ \bm{\zeta} \bm{\epsilon}} \left( \bm{\rho} \right) - \mathbf{R}_{\bm{\zeta} \bm{\eta}} \left( \bm{\rho} \right) \bm{\theta}  =  \mathbf{J} \left( \bm{\theta} \right) \bm{\rho}, \\
    \mathbf{R} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau-1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{\bm{\phi}}^\top \left( \tau \right) , \\     
    \mathbf{q} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau-1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{y} \left( \tau \right) \\     
    \mathbf{b} \left( t \right)  &=  \mathbf{q} \left( t \right) - \mathbf{h} \left( \bm{\rho}, \ \bm{\theta} \right)
    , \end{aligned} \end{equation}
    then the equation (\ref{eqn:GIVE}) is equivalent to
    \begin{equation} \tag{4.35} \mathbf{R} \left( t \right) \bm{\theta} = \mathbf{b} \left( t \right) . \end{equation}
    This representation allows a recursive form
    \begin{equation} \tag{4.36} \begin{aligned} 
    \mathbf{R} \left( t+1 \right)  &= \mathbf{R} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{\bm{\phi}}^\top \left( t+1 \right) - \mathbf{R} \left( t \right) \right), \\
    \mathbf{b} \left( t+1 \right)  &= \mathbf{b} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{y} \left( t+1 \right) - \mathbf{b} \left( t \right) - \mathbf{h} \left( \bm{\rho}, \ \bm{\theta} \right) \right), \\
    \widehat{\bm{\theta}} \left( t+1 \right)  &=  \left( \mathbf{R}^\top \left( t+1 \right) \mathbf{R} \left( t+1 \right) \right)^{-1} \mathbf{R}^\top \left( t+1 \right) \mathbf{b} \left( t+1 \right) \\
    \mathbf{q} \left( t+1 \right)  &= \mathbf{q} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{y} \left( t+1 \right) - \mathbf{q} \left( t \right) \right) \\
    \widehat{\bm{\rho}} \left( t+1 \right)  &=  \mathbf{J}^\dagger \left( \widehat{\bm{\theta}} \left( t \right) \right) \left( \mathbf{q} \left( t \right) - \mathbf{R} \left( t \right) \widehat{\bm{\theta}} \left( t \right) \right) , \end{aligned} \end{equation}
    where $\mathbf{J}^\dagger \left( \widehat{\bm{\theta}} \left( t \right) \right)$ is the pseudo-inverse of $\mathbf{J} \left( \bm{\theta} \right)$.

@@

    \item  Why do you always study the statistical properties of your least squares solution with high SNR only? What do you get for SNR = 10 dB for instance? 
    \item Give the values of the matrices you use for simulating data sets so that the reader can test what you have carried out.

@@

    \renewcommand{\thefigure}{2.2}
    \begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterIntroduction/fig/Fig_2.pdf} 
    \caption{ \label{fig:thermometer} 
    \color{blue} This simulation shows an example of the temperature change experimented by a thermometer, with $k_T=0.5 s^{-1}$, caused by applying a step input of 67 $\circ \mathrm{C}$  when the thermometer was initially at 23 $\circ \mathrm{C}$. In this example, we should wait 10 s to have an accurate estimation of the input. \color{black} }
    \end{figure}

    \renewcommand{\thefigure}{2.3}
    \begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterIntroduction/fig/Fig_3.pdf} 
    \caption{ \label{fig:scale} 
    \color{blue} This simulation illustrate the displacement observed when a constant input $u=1 \mathrm{kg}$ is applied to a weighing sensor in equilibrium. The sensor parameters are $k_{\mathrm{d}} = 5 \mathrm{N s/m}$, and $k_{\mathrm{s}} = 10000 \mathrm{N/m}$. In this conditions, the user has to wait more that 2 s to observe a stabilization on the scale and to estimate the mass from reading the final value of the displacement. \color{black}  }
    \end{figure}

>>>>

    \renewcommand{\thefigure}{6.8}
    \begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterRampInput/fig/Fig_7.pdf} 
    \caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor's initial conditions are estimated with the ML method. After three iterations, \color{blue} the average of the relative error is smaller than 5 \% for each and every of the estimates. The estimate $\widehat{x}_{\mathrm{ini,2}}$ has the larger relative error near to 1 \%. \color{black} }
    \end{figure}

@@

	\color{blue}
	The normality assumption is necessary to provide more prior information into the method formulation.
    The second order Taylor series expansion (4.4) is developed considering the perturbation noise that gets into the regression matrix from the sensor transient response.
    Assuming only that the perturbation noise is distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$ is not enough.
    If we additionally assume normality for the perturbation noise, then we benefit from the knowledge of the third moment equal to zero due to symmetry, and the fourth moment equal to three times the squared variance\color{black}.

@@

\begin{itemize}
	\item Two practical questions. On page 35-49 you consider a weight of 138.x grams. You sample relatively slowly (4 kHz) and after 500 samples we can consider that we have correct estimation. If I have a container that fills up and weight is constantly changing. Can we tackle this?

	{\bfseries Yes, that is exactly the idea considered by the affine input estimation method. Since the data-driven estimation method can track the changes of the input, the parameters of an input model that describes the constant variation can be recovered if we add the forgetting factor to the recursive solution algorithm. }

	\item  Consider an optical fiber link with a detector capable of 100 million measurements per second. Can I speed up the link by using your method? You have to do calculations: how high can you go in speed before running into troubles with DSP?

	{\bfseries The optical fiber link transmits digital pulses. Since the period would be 1 ns, the step input method implementation would require to sample in much shorter time. Assuming that the duty cyce is 50 \%, and that the number of samples required for the step detection are 200, then the sampling time should be 1.25 fs or lower. This could be a constraint for the commutation circuits and we would be reaching a feasibility limit. However, theoretically the method would perform detecting the pulses level changes.  }

@@

    The errors-in-variables problem formulation (3.18) of the data-driven step input estimation method can be converted into an instrumental-variables problem.
    A guide to do this conversion is found in \cite{Soderstrom18} and in \cite{Pintelon12Book}. 
    The instrumental variable $\widetilde{\mathbf{Z}}$ has to be selected so that the normal equations (3.19) are expresed instead as
    \begin{equation} \tag{4.22} \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{y}} . \label{eqn:neq_siv} \end{equation}
    The selection of the instrumental variable has to be done in such a way that is correlated with the regression matrix $\widetilde{\mathbf{K}}$, and uncorrelated with the perturbations $\bm{\epsilon}$.
    Assuming that the block-Hankel matrix $\widetilde{\mathbf{K}}$ is given by
    \begin{equation} \tag{4.23} \widetilde{\mathbf{K}} = \begin{bmatrix} G & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ G & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(n+1) & \Delta \widetilde{y}(n+2) & \cdots & \Delta \widetilde{y}(2n) \end{bmatrix} , \label{eqn:matrixK_r} \end{equation}
    then, the instrumental variable $\widetilde{\mathbf{Z}}$ should not contain the elements that construct $\widetilde{\mathbf{K}}$, i.e., the elements of the vector $\left( \widetilde{y}(1), \ldots, \widetilde{y}(2n+1) \right)^\top$.
    A possibility to generalize this idea is expressing the instrumental variable as
    \begin{equation} \tag{4.24} \widetilde{\mathbf{Z}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-4n) & \Delta \widetilde{y}(t-4n+1) & \cdots & \Delta \widetilde{y}(t-3n-1) \\ G & \Delta \widetilde{y}(t-4n+1) & \Delta \widetilde{y}(t-4n+2) & \cdots & \Delta \widetilde{y}(t-3n) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-3n) & \Delta \widetilde{y}(t-3n+1) & \cdots & \Delta \widetilde{y}(t-2n-1) \end{bmatrix} , \label{eqn:matrixZ_t} \end{equation}
    for a regression matrix
    \begin{equation} \tag{4.25} \widetilde{\mathbf{K}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-2n+1) & \Delta \widetilde{y}(t-2n+2) & \cdots & \Delta \widetilde{y}(t-n) \\ G & \Delta \widetilde{y}(t-2n+2) & \Delta \widetilde{y}(t-2n+3) & \cdots & \Delta \widetilde{y}(t-n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-n+1) & \Delta \widetilde{y}(t-n+2) & \cdots & \Delta \widetilde{y}(t) \end{bmatrix} , \label{eqn:matrixK_t} \end{equation}
    for $t = 4n+1, \ldots, N$.

@@

    \subsubsection*{4.1.4.2 Recursive generalized instrumental variables method}
    If we define
    \begin{equation} \tag{4.34} \begin{aligned} \mathbf{h} \left( \bm{\rho}, \bm{\theta} \right)  &=  \mathbf{r}_{ \bm{\zeta} \bm{\epsilon}} \left( \bm{\rho} \right) - \mathbf{R}_{\bm{\zeta} \bm{\eta}} \left( \bm{\rho} \right) \bm{\theta}  =  \mathbf{J} \left( \bm{\theta} \right) \bm{\rho}, \\
    \mathbf{R} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau=1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{\bm{\phi}}^\top \left( \tau \right) , \\     
    \mathbf{q} \left( t \right)  &=  \dfrac{1}{t} \sum_{\tau=1}^{t} \widetilde{\mathbf{z}} \left( \tau \right) \widetilde{y} \left( \tau \right) \\     
    \mathbf{b} \left( t \right)  &=  \mathbf{q} \left( t \right) - \mathbf{h} \left( \bm{\rho}, \ \bm{\theta} \right)
    , \end{aligned} \end{equation}
    then the equation (\ref{eqn:GIVE}) is equivalent to
    \begin{equation} \tag{4.35} \mathbf{R} \left( t \right) \bm{\theta} = \mathbf{b} \left( t \right) . \end{equation}
    This representation allows a recursive form
    \begin{equation} \tag{4.36} \begin{aligned} 
    \mathbf{R} \left( t+1 \right)  &= \mathbf{R} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{\bm{\phi}}^\top \left( t+1 \right) - \mathbf{R} \left( t \right) \right), \\
    \mathbf{b} \left( t+1 \right)  &= \mathbf{b} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{y} \left( t+1 \right) - \mathbf{b} \left( t \right) - \mathbf{h} \left( \bm{\rho}, \ \bm{\theta} \right) \right), \\
    \widehat{\bm{\theta}} \left( t+1 \right)  &=  \left( \mathbf{R}^\top \left( t+1 \right) \mathbf{R} \left( t+1 \right) \right)^{-1} \mathbf{R}^\top \left( t+1 \right) \mathbf{b} \left( t+1 \right) \\
    \mathbf{q} \left( t+1 \right)  &= \mathbf{q} \left( t \right) + \dfrac{1}{t+1} \left( \widetilde{\mathbf{z}} \left( t+1 \right) \widetilde{y} \left( t+1 \right) - \mathbf{q} \left( t \right) \right) \\
    \widehat{\bm{\rho}} \left( t+1 \right)  &=  \mathbf{J}^\dagger \left( \widehat{\bm{\theta}} \left( t \right) \right) \left( \mathbf{q} \left( t \right) - \mathbf{R} \left( t \right) \widehat{\bm{\theta}} \left( t \right) \right) , \end{aligned} \end{equation}
    where $\mathbf{J}^\dagger \left( \widehat{\bm{\theta}} \left( t \right) \right)$ is the pseudo-inverse of $\mathbf{J} \left( \bm{\theta} \right)$.

@@

    \item  Why do you always study the statistical properties of your least squares solution with high SNR only? What do you get for SNR = 10 dB for instance?

    {\bfseries The second Taylor series expansion of the inverse matrix $(\mathbf{I} + \mathbf{M})^{-1}$ is valid when the spectral radius of $\mathbf{M}$ is lower than 1. The spectral radius depends on the perturbation noise variance. As the noise variance increases, for fixed exact data, the SNR decreases and the spectral radius grows, as well as the error introduced by the Taylor series expansion. We observe that for large SNR the Taylor series expansion allows us to use the derived expressions that calculate the bias and the variance. When the relative error of these preditions increase, below 40 dB for example, the spectral radius is getting closer and eventually going over 1. With SNR = 10 dB the Taylor series expansion of the inverse matrix introduces an error unacceptably large. }

    \item Give the values of the matrices you use for simulating data sets so that the reader can test what you have carried out.

@@

    \renewcommand{\thefigure}{2.2}
    \begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterIntroduction/fig/Fig_2.pdf} 
    \caption{ \label{fig:thermometer} 
    \color{blue} This simulation shows an example of the temperature change experimented by a thermometer, with $k_T=0.5 s^{-1}$, caused by applying a step input of 67 $\circ \mathrm{C}$  when the thermometer was initially at 23 $\circ \mathrm{C}$. In this example, we should wait 10 s to have an estimation of the input with a relative error lower than 0.5 \%. \color{black} }
    \end{figure}

    \renewcommand{\thefigure}{2.3}
    \begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterIntroduction/fig/Fig_3.pdf} 
    \caption{ \label{fig:scale} 
    \color{blue} This simulation illustrates the displacement observed when a constant input $u=1 \mathrm{kg}$ is applied to a weighing sensor in equilibrium. The sensor parameters are $k_{\mathrm{d}} = 5 \mathrm{N s/m}$, and $k_{\mathrm{s}} = 10000 \mathrm{N/m}$. In this conditions, the user has to wait more that 2 s to observe a stabilization on the scale and to estimate the mass from reading the final value of the displacement with a relative error smaller than 0.72 \%. \color{black}  }
    \end{figure}
